Inattentive responding can induce spurious associations between task behavior and symptom measures

### 2023-04-10 manuscripts/main.tex
checkpointing manuscript progress

@article{hunter2019excessive,
  title={Excessive deliberation in social anxiety},
  author={Hunter, Lindsay E and Meer, Elana A and Gillan, Claire M and Hsu, Ming and Daw, Nathaniel D},
  journal={bioRxiv},
  pages={522433},
  year={2019},
  publisher={Cold Spring Harbor Laboratory}
}

@@

With the advantages of online data collection also come specific drawbacks. Since participants recruited from online labor markets are typically completing experiments in their homes, they may be more likely to be distracted or multi-tasking during an experiment. They may also be more likely to use heuristic response strategies with the intention to minimise expenditure of time and cognitive effort (e.g., responding randomly on self-report surveys or behavioral tasks). Here, we will refer to such inattentive or low-effort behaviors as careless/insufficient effort (C/IE) responding \cite{huang2012detecting, curran2016methods}. Among researchers using online labor markets, a common view is that poor-quality data resulting from C/IE responding can simply be treated as a source of unsystematic measurement error that can be overcome with increased sample sizes \cite{gillan2016taking, chandler2020participant}. Common practice in online behavioral research is to mitigate poor-quality data using the same screening methods that are typically used in in-person data collection (e.g., excluding participants who perform at- or below-chance on behavioral tasks). However, these methods may be specifically inappropriate for online psychiatry studies, as we detail below.

Here we wish to draw special attention to an underappreciated feature of psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom distributions tend to be heavily positively skewed \cite{lowe2008validation, tomitaka2018distributional}. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey attentively \cite{chandler2020participant, ophir2020turker, king2018random}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient attention or effort.

When participants complete both symptom surveys and behavioral tasks---a common study design in computational psychiatry---this artifact has the potential to induce spurious correlations between symptom self-reports and task behavior. That is, while C/IE behavior is traditionally thought of as a source of noise that can result in type II (false negative) errors, here we suggest that in large-scale online psychiatric studies it can instead result in type I (false positive) errors. Concretely, if the same participants who engage in C/IE responding on surveys (and who therefore inaccurately report high levels of psychiatric symptoms) also respond with insufficient effort on behavioral tasks, this can cause experimenters to observe an entirely spurious correlation between greater symptom severity and worse task performance (see Figure \ref{fig:simulation}). A similar effect has been well documented in personality psychology, where the presence of C/IE responding can induce correlations between questionnaires, and bias estimated factors in factor analysis \cite{huang2012detecting, robinson2014inaccurate, huang2015insufficient, chandler2020participant, arias2020little}.

Here, we demonstrate the real risk that C/IE responding can lead to spurious symptom-task correlations in computational psychiatry research. First, we asked to what extent recent studies in computational psychiatry screen participants based on self-report symptom data. We found that the majority of these studies did not screen participants' survey data at all, and that very few followed best-practice recommendations for survey data screening. We then asked whether behavioral screening alone was sufficient to identify participants engaging in C/IE responding on psychiatric symptom surveys. In two new datasets from two separate online labor marketplaces, we found that screening based on task behavior fails to completely identify participants engaging in C/IE responding on surveys. Lastly, we investigated whether, under these circumstances, C/IE responding led to spurious correlations between symptom severity and task performance for positively-skewed symptom measures. Consistent with the logic set out above, we confirmed that failure to appropriately screen out C/IE survey responding in the proof-of-concept datasets that we collected would have produced a number of spurious correlations between task behavior and self-reported symptoms that are abolished when data are screened more thoroughly.

@@

Among studies that we reviewed, approximately 80\% (39/49) used at least one method to identify C/IE responding in task behavior (see Table \ref{tab:tab01}). Of these, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was statistically indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

@@

To measure the correspondence of screening measures estimated from task and self-report behavior, we conducted an online behavioral experiment involving a simple decision-making task and a battery of commonly used self-report psychiatric symptom measures (see Methods). A final sample of 386 participants from the Amazon Mechanical Turk (N=186) and Prolific (N=200) online labor markets completed a probabilistic reversal-learning task and 5 self-report symptom measures. The reversal-learning task required participants to learn through trial-and-error which of three options yielded reward most often, and was modeled after similar tasks used to probe reinforcement-learning deficits in psychiatric disorders \cite{waltz2007probabilistic, mukherjee_reward_2020}. The five self-report measures were the 7-up (which measures symptoms of hypomania), the 7-down (which measures symptoms of depression), the GAD-7, (which measures generalized anxiety symptoms), the BIS/BAS (which measures reward and punishment motivations), the SHAPS (which measures anhedonia symptoms), and the PSWQ (which measures worry symptoms), and were chosen based on previous literature to have a variety of expected response distributions (symmetric and asymmetric). In line with current best-practice recommendations in personality psychology \cite{huang2015detecting}, each self-report instrument included one `infrequency' item that could be used to identify C/IE responses in survey data (see Methods for a list of infrequency items). The entire experiment (surveys and task) was designed to require 10 minutes on average to complete (observed mean = 10.28 minutes). To minimize any influence of fatigue on survey responding, participants completed the surveys prior to beginning the task.

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low; that is, comparably higher scores are more likely for C/IE participants responding at random on a right skewed questionnaire. In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; see Figure \ref{fig:fig01}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistically significant for the majority of symptom measures (Table \ref{tab:tab02}).

Hereafter, we use the infrequency-item method as a primary means of identifying C/IE responding in our data. To verify this approach, we conducted three validation analyses. The first analysis compared estimated internal consistency of self-report measures between the C/IE and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the C/IE responding group \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}. In line with this reasoning, we observed a reduction in Cronbach's $\alpha$ in the C/IE group for the majority of survey instruments (Table \ref{tab:tab02}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance given the difference in participant numbers between groups (two-tailed, paired-samples $t$-test: $t(6)=3.689$, $p=0.021$, $d=1.506$, 95\% CI = [0.048, 0.141]).

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = 0.67, 95\% \ \text{HDI} = [-0.78, -0.55]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less extreme} symptoms or personality traits on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms or personality traits) despite these inventories measuring strongly correlated constructs (i.e., depression and anhedonia, anxiety and behavioral inhibition).

@@

Second, we used the Dice similarity coefficient to quantify agreement between different screening methods in the set of participants flagged for exclusion (Figure \ref{fig:fig02}, right panel). This approach quantifies the degree of overlap between the set of would-be excluded participants based on different screening measures under a common exclusion rate. Though some measures have relatively clear threshold cutoffs (e.g., chance level performance for task accuracy), the majority of the measures evaluated here do not. As such, we evaluated the measures with respect to the top 10\% of ``suspect" participants flagged by each measure, corresponding roughly to the fraction of participants having performed at chance levels on the reversal-learning task. (Results of the same analysis repeated for the top 25\% of ``suspicious" participants --- corresponding roughly to the fraction of participants flagged by the infrequency-item measure --- produced similar results; see Table SXX.) Results were largely consistent with the correlation analysis: few pairs of task and self-report screening measures achieved levels of agreement greater than what would be expected by chance. The only significant cross-modality pair identified --- between the infrequency item and choice accuracy measures --- has a similarly coefficient less than 0.4. In other words, when these two measures are used to identify the top 10\% of participants most strongly suspected of C/IE responding, they agree on only two out of every five participants. Screening on choice accuracy alone (the most common method identified in our literature review) would fail to identify the majority of participants most likely engaging in C/IE responding as determined by the infrequency items.

@@

When no rejections based on C/IE responding was applied (i.e., all participants were included in the analysis; Figure \ref{fig:fig03}A), many significant correlations emerged between measures of task behavior and symptom scores, in particular for 4 of the self-report instruments (7-up, which measures symptoms of hypomania; 7-down, which measures symptoms of depression; GAD-7, which measures generalized anxiety symptoms; and BIS, which measures tendencies related to behavioral inhibition). Consistent with our predictions, the majority of these correlations involved symptom measures with asymmetric total score distributions. Attending to only the most skewed measures (i.e., 7-up, 7-down, GAD-7), symptom endorsement was correlated with almost every behavioral measure. That is, significant correlations were not restricted only to general behavioral measures often used as proxies for participant effort (e.g., accuracy, inverse temperature $\beta$) but also to measures of specific theoretical interest, such as asymmetry of learning from positive and negative reward prediction errors ($\kappa$). Conversely, we found few significant correlations among symptom measures with more symmetric distributions. This is despite the fact these scales measure similar symptoms and syndromes (e.g., anxiety as measured by the GAD-7 and worry as measured by the PSWQ; depression as measured by the 7-down and anhedonia as measured by the SHAPS).

@@

In stark contrast, when we rejected participants based on self-report screening  (removing 22\% of participants who endorsed one or more invalid or improbable responses on the infrequency items; Figure \ref{fig:fig03}C), the number of significant correlations was markedly reduced, particularly for several of the most skewed symptom measures (7-down, GAD-7) and proxy measures of task attentiveness (e.g., accuracy, inverse temperature). This pattern of correlations was largely similar when rejections were applied based on both task and self-report screening measures (Figure \ref{fig:fig03}D). We also note that with stricter screening, the remaining significant correlations were, for the most part (but not always), weaker.

These findings suggest that many of the significant behavior-symptom correlations observed without strict participant screening may indeed be spurious correlations driven by C/IE responding. Importantly, screening based on task behavior alone did not adequately protect against spurious symptom-behavior correlations in the presence of skewed distributions of symptom endorsement. For instance, consider the 7-down scale, a measure of trait depression: had we not screened participants based on infrequency items, we would have erroneously concluded that there were many significant associations between reversal-learning task performance and self-reported depression. Screening on self-report data allowed us to identify that each of these depression-behavior correlations was likely to be entirely spurious.

One possible objection to this interpretation is that the reduction in significant correlations following self-report screening was a result of the reduced sample size after removal of C/IE respondents (which comprised over 20\% of the sample). To test this alternative hypothesis, we performed the same correlation analysis after removing random subsets of participants, fixing the sample size to that obtained after excluding C/IE respondents. In this case, the pattern of significant correlations was more similar to that before screening than after screening using the infrequency measure (two-tailed, paired-samples $t$-test: $t = 262.490$, $p < 0.001$, $d = 3.713$, 95\% CI = [0.136, 0.138]; Figure S2, compare to Figure \ref{fig:fig03}A). Thus, the reduction in significant correlations following screening was unlikely to be driven solely by a reduction in statistical power.

@@

For the sake of brevity, we report here only the main pattern of findings (all results are reported in Appendix B). In the replication sample, 55 out of 393 participants (14\%) endorsed a logically invalid or improbable response on one or more of the infrequency items when completing the self-report measures. This is roughly two-thirds of the number of participants who were flagged for C/IE responding in the original study, suggesting that the newer quality assurance protocols used by the online platforms are at least partially effective.

In the self-report symptom measures, we replicated the finding that total scores were noticeably exaggerated in participants suspected of C/IE responding, but only for symptom measures where overall rates of symptom endorsement were the lowest. Similarly, we again found that task-based screening and self-report screening measures showed low correspondence; that is, excluding participants on the basis of poor behavioral performance would not have identified and removed participants who engaged in C/IE responding on self-report surveys.

Finally, when we did not apply any exclusions, we observed spurious correlations between performance on the two-step task and total scores for both symptom and personality self-report measures with a mean-shift in scores between attentive participants and participants suspected of C/IE responding. In contrast with our original findings, however, we found that excluding participants based on self-report \textit{or} task screening measures was sufficient to ablate these spurious correlations.

@@

To explore this possibility, we embedded attention checks into the self-report measures of two studies of patients with major depressive disorder (see Appendix C for details). Specifically, N=45 psychiatric patients (confirmed to meet criteria for a diagnosis of major depressive disorder through a structured clinical interview) and N=20 healthy controls, all recruited through the Rutgers-Princeton Center for Computational Cognitive Neuropsychiatry (i.e., not on online labor platforms), completed a series of self-report symptom measures, online, on their computers from the comfort of their homes. In total, 16 of 65 (24.6\%) participants failed one or more attention checks. Subdivided by group, 6 of 20 (30\%) healthy participants and 10 of 45 (22\%) psychiatric participants were flagged for C/IE responding.

Using these data, we computed pairwise Bayes factors comparing three candidate models: attention check failure rates are equal between healthy and psychiatric participants (M1); failure rates are greater in psychiatric patients (M2); and failure rates are greater in healthy participants (M3). The model assuming equal rates of failure between healthy and psychiatric participants was 2.88 times more likely than the model assuming greater rates for psychiatric patients. In turn, the model assuming lower rates of failure for psychiatric patients was 1.27 times more likely than the model assuming equal rates. Finally, the model assuming lower rates of failure for psychiatric patients was 3.65 times more likely than the model assuming higher rates for psychiatric patients. Only the final comparison exceeds the cutoff value of 3, which is conventionally treated as the minimal amount of evidence required to treat a difference in model fit as meaningful. Although the size of the sample precludes any definitive conclusion, it is noteworthy that the model least consistent with the data was the one where psychiatric patients are more likely to fail infrequency item attention checks. These data suggest, therefore, that it is unlikely that individuals with high psychiatric symptom severity were disproportionately flagged for C/IE responding in the main analyses. Accordingly, we conclude that the screening measures we are suggesting are not likely to result in overcontrol bias and false-negative correlations between tasks and symptom measures.

@@

In this manuscript, we highlighted a particular set of circumstances, common in computational psychiatry research done on large online samples, in which spurious correlations may arise between task behavior and self-reported symptomology. When the ground-truth prevalence of a symptom is low in the general population, participants who respond carelessly on measures assessing this symptom may erroneously appear as symptomatic. A less careful pattern of responding on tasks used to measure cognitive constructs can then masquerade as a correlation between alteration in these constructs and symptom dimensions. We found repeated evidence for this pernicious pattern in participants recruited from two popular online labor platforms. False-positive rates for these spurious correlations \emph{increased} with sample size, because the correlations are due to measurement bias, not measurement noise. Importantly, we found that screening on task behavior alone was insufficient to identify participants engaging in C/IE responding and prevent some false-positive correlations. Unfortunately, a literature review identified this type of screening as the most common practice in online computational psychiatry studies. We recommend instead to screen and exclude participants based on responding on surveys, a practice that abolished many spurious behavior-symptom correlations in our data.

@@

Viewed from the perspective of effort expenditure, our results suggest that participants appraised the cost/benefit trade-off differently for behavioral tasks and self-report surveys. Specifically, we found that only 7\% of participants in the first study were at chance-level performance in the task, compared to more than 22\% of participants who failed one or more attention-check items in the self-report surveys (a finding that qualitatively replicated in a second study involving a different task). Moreover, different measures of C/IE responding were weakly or not at all correlated between task behavior and self-report responses. This suggests the motivation for effortful responding was greater in the behavioral task, though precisely why is unclear. One possibility is that we gave participants a monetary incentive for attentive responding only during the task (a common practice, according to our literature review). A second possibility is that participants expected fewer consequences for C/IE responding during the self-report surveys, a reasonable assumption in light of how infrequently previous experiments have screened self-report data. Alternatively, participants may have found the gamified behavioral task more engaging or the self-report inventory more tedious. Regardless of the reason, this discrepancy reinforces our observations concerning the inadequacy of behavioral-task screening as a stand-alone method for identifying C/IE responding. Since, in general, participants may appraise costs and benefits of effortful responding differently for behavioral tasks and self-report surveys, screening for C/IE responding on one data modality may in general be unsuitable for identifying it in the other. We therefore recommend screening on each component of an experiment.

>>>>

@article{hunter2022increased,
  title={Increased and biased deliberation in social anxiety},
  author={Hunter, Lindsay E and Meer, Elana A and Gillan, Claire M and Hsu, Ming and Daw, Nathaniel D},
  journal={Nature Human Behaviour},
  volume={6},
  number={1},
  pages={146--154},
  year={2022},
  publisher={Nature Publishing Group UK London}
}

@@

With the advantages of online data collection also come specific drawbacks. Since participants recruited from online labor markets are typically completing experiments in their homes, they may be more likely to be distracted or multi-tasking during an experiment. They may also be more likely to use heuristic response strategies with the intention to minimize expenditure of time and cognitive effort (e.g., responding randomly on self-report surveys or behavioral tasks). Here, we will refer to such inattentive or low-effort behaviors as careless/insufficient effort (C/IE) responding \cite{huang2012detecting, curran2016methods}. Among researchers using online labor markets, a common view is that poor-quality data resulting from C/IE responding can simply be treated as a source of unsystematic measurement error that can be overcome with increased sample sizes \cite{gillan2016taking, chandler2020participant}. Common practice in online behavioral research is to mitigate poor-quality data using the same screening methods that are typically used in in-person data collection (e.g., excluding participants who perform at- or below-chance on behavioral tasks). However, these methods may be specifically inappropriate for online psychiatry studies, as we detail below.

Here we wish to draw special attention to an underappreciated feature of psychiatric research using self-report symptom surveys. In such surveys, participants rate their endorsement of various psychiatric symptoms and, since most individuals in the general population tend to endorse no or few symptoms in many symptom domains, the resulting ground-truth symptom score distributions tend to be heavily positively skewed \cite{lowe2008validation, tomitaka2018distributional}. In this situation, the assumption that C/IE responding merely increases unsystematic measurement noise becomes untenable. Because of the positive skew in the ground-truth symptom distribution, participants who respond carelessly to the symptom survey are more likely to report higher levels of symptom endorsement relative to participants who complete the survey attentively \cite{chandler2020participant, ophir2020turker, king2018random}. Consequently, unless C/IE survey responses are carefully identified and removed, a considerable proportion of putatively symptomatic individuals in an online sample may, in fact, be participants who have not engaged with the experiment with sufficient attention or effort.

When participants complete both symptom surveys and behavioral tasks---a common study design in computational psychiatry---this artifact has the potential to induce spurious correlations between symptom self-report scores and task behavior. That is, while C/IE behavior is traditionally thought of as a source of noise that can result in type II (false negative) errors, here we suggest that in large-scale online psychiatric studies it can instead result in type I (false positive) errors. Concretely, if the same participants who engage in C/IE responding on surveys (and who therefore inaccurately report high levels of psychiatric symptoms) also respond with insufficient effort on behavioral tasks, this can cause experimenters to observe an entirely spurious correlation between greater symptom severity and worse task performance (see Figure \ref{fig:simulation}). A similar effect has been well documented in personality psychology, where the presence of C/IE responding can induce correlations between questionnaires, and bias factor estimation in factor analysis \cite{huang2012detecting, robinson2014inaccurate, huang2015insufficient, chandler2020participant, arias2020little}.

Here, we demonstrate the real risk that C/IE responding can lead to spurious symptom-task correlations in computational psychiatry research. First, we asked to what extent recent studies in computational psychiatry screen participants based on self-report symptom data. We found that the majority of these studies did not screen participants' survey data at all, and that very few followed best-practice recommendations for survey data screening. We then asked whether behavioral screening alone was sufficient to identify participants engaging in C/IE responding on psychiatric symptom surveys. In two new datasets from two separate online labor markets, we found that screening based on task behavior fails to completely identify participants engaging in C/IE responding on surveys. Lastly, we investigated whether, under these circumstances, C/IE responding led to spurious correlations between symptom severity and task performance for positively-skewed symptom measures. Consistent with the logic set out above, we confirmed that failure to appropriately screen out C/IE survey responding in the proof-of-concept datasets that we collected would have produced a number of spurious correlations between task behavior and self-reported symptoms that are abolished when data are screened more thoroughly.

@@

Among studies that we reviewed, approximately 80\% (39/49) used at least one method to identify C/IE responding in task behavior (Table \ref{tab:tab01}). Of these, just over half relied on a single screening method, with considerable heterogeneity in behavior screening methods across studies. Most common (46\% of all studies) was identifying participants whose performance was statistically indistinguishable from chance-level on some measure of accuracy. Almost as common (38\%) was screening based on low response variability (i.e., excluding participants who predominantly responded in the same fashion across trials, such as using only a single response key).

@@

To measure the correspondence of screening measures estimated from task and self-report behavior, we conducted an online behavioral experiment involving a simple decision-making task and a battery of commonly used self-report psychiatric symptom measures (see Methods). A final sample of 386 participants from the Amazon Mechanical Turk (N=186) and Prolific (N=200) online labor markets completed a probabilistic reversal-learning task and 5 self-report symptom measures. The reversal-learning task required participants to learn through trial-and-error which of three options yielded reward most often, and was modeled after similar tasks used to probe reinforcement-learning deficits in psychiatric disorders \cite{waltz2007probabilistic, mukherjee_reward_2020}. The five self-report measures were the 7-up (which measures symptoms of hypomania), the 7-down (which measures symptoms of depression), the GAD-7, (which measures generalized anxiety symptoms), the BIS/BAS (which measures reward and punishment motivations), the SHAPS (which measures anhedonia symptoms), and the PSWQ (which measures worry symptoms), and were chosen based on previous literature to have a variety of expected response distributions (symmetric and asymmetric). In line with current best-practice recommendations \cite{huang2015detecting}, each self-report instrument included one `infrequency' item that could be used to identify C/IE responses in survey data (see Methods for a list of infrequency items). The entire experiment (surveys and task) was designed to require 10 minutes on average to complete (observed mean = 10.28 minutes). To minimize any influence of fatigue on survey responding, participants completed the surveys prior to beginning the task.

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low; that is, comparably higher scores are more likely for C/IE participants responding at random on a questionnaire with a right-skewed score distribution. In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; Figure \ref{fig:fig01}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistically significant for the majority of symptom measures (Table \ref{tab:tab02}).

Hereafter, we use the infrequency-item method as a primary means of identifying C/IE responding in our data. To verify this approach, we conducted three validation analyses. The first analysis compared estimated internal consistency of self-report measures between the C/IE and attentive groups. The logic is that, if C/IE responding manifests as a tendency to respond randomly, we should expect to see a decrease in the consistency of a measure in the C/IE responding group \cite{maniaci2014caring, desimone2018dirty, desimone2018differential}. In line with this reasoning, we observed a reduction in Cronbach's $\alpha$ in the C/IE group for the majority of survey instruments (Table \ref{tab:tab02}). A permutation test confirmed that the average decrease in internal consistency across measures was greater than would be expected by chance given the difference in participant numbers between groups (two-tailed, paired-samples $t$-test: $t(6)=-3.689$, $p=0.021$, $d=1.506$, 95\% CI = [-0.048, -0.141]).

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = -0.67, 95\% \ \text{HDI} = [-0.78, -0.55]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less extreme} symptoms or personality traits on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms or personality traits) despite these inventories measuring strongly correlated constructs (i.e., depression and anhedonia, anxiety and behavioral inhibition).

@@

Second, we used the Dice similarity coefficient to quantify agreement between different screening methods in the set of participants flagged for exclusion (Figure \ref{fig:fig02}, right panel). This approach quantifies the degree of overlap between the set of would-be excluded participants based on different screening measures under a common exclusion rate. Though some measures have relatively clear threshold cutoffs (e.g., chance level performance for task accuracy), the majority of the measures evaluated here do not. As such, we evaluated the measures with respect to the top 10\% of ``suspect" participants flagged by each measure, corresponding roughly to the fraction of participants having performed at chance levels on the reversal-learning task. (Results of the same analysis repeated for the top 25\% of ``suspicious" participants --- corresponding roughly to the fraction of participants flagged by the infrequency-item measure --- produced similar results; see Table S5.) Results were largely consistent with the correlation analysis: few pairs of task and self-report screening measures achieved levels of agreement greater than what would be expected by chance. The only significant cross-modality pair identified --- between the infrequency item and choice accuracy measures --- has a Dice similarity coefficient less than 0.4. In other words, when these two measures are used to identify the top 10\% of participants most strongly suspected of C/IE responding, they agree on only two out of every five participants. Screening on choice accuracy alone (the most common method identified in our literature review) would fail to identify the majority of participants most likely engaging in C/IE responding as determined by the infrequency items.

@@

When no rejections based on C/IE responding was applied (i.e., all participants were included in the analysis; Figure \ref{fig:fig03}A), many significant correlations emerged between measures of task behavior and symptom scores, in particular for 4 of the self-report instruments (7-up, which measures symptoms of hypomania; 7-down, which measures symptoms of depression; GAD-7, which measures generalized anxiety symptoms; and BIS, which measures tendencies related to behavioral inhibition). Consistent with our predictions, the majority of these correlations involved symptom measures with asymmetric score distributions. Attending to only the most skewed measures (i.e., 7-up, 7-down, GAD-7), symptom endorsement was correlated with almost every behavioral measure. That is, significant correlations were not restricted only to general behavioral measures often used as proxies for participant effort (e.g., accuracy, inverse temperature $\beta$) but also to measures of specific theoretical interest, such as asymmetry of learning from positive and negative reward prediction errors ($\kappa$). Conversely, we found few significant correlations among symptom measures with more symmetric distributions. This is despite the fact these scales measure similar symptoms and syndromes (e.g., anxiety as measured by the GAD-7 and worry as measured by the PSWQ; depression as measured by the 7-down and anhedonia as measured by the SHAPS).

@@

In stark contrast, when we rejected participants based on self-report screening  (removing 22\% of participants who endorsed one or more invalid or improbable responses on the infrequency items; Figure \ref{fig:fig03}C), the number of significant correlations was markedly reduced, particularly for several of the most skewed symptom measures (7-down, GAD-7) and proxy measures of task attentiveness (e.g., accuracy, inverse temperature). This pattern of correlations was largely similar when rejections were applied based on both task and self-report screening measures (Figure \ref{fig:fig03}D). We also note that with stricter screening, the remaining significant correlations were, for the mostly but not always, weaker (Tables S6--S9).

These findings suggest that many of the significant behavior-symptom correlations observed without strict participant screening may indeed be spurious correlations driven by C/IE responding. Importantly, screening based on task behavior alone did not adequately protect against spurious symptom-behavior correlations in the presence of skewed distributions of symptom endorsement. For instance, consider the 7-down scale, a measure of trait depression: had we not screened participants based on infrequency items, we would have erroneously concluded that there were many significant associations between reversal-learning task performance and self-reported depression. Screening on self-report data allowed us to identify that each of these depression-behavior correlations was likely to be spurious.

One possible objection to this interpretation is that the reduction in significant correlations following self-report screening was a result of the reduced sample size after removal of C/IE respondents (which comprised over 20\% of the sample). To test this alternative hypothesis, we performed the same correlation analysis after removing random subsets of participants, fixing the sample size to that obtained after excluding C/IE respondents. In this case, the pattern of significant correlations was more similar to that before screening than after screening using the infrequency measure (two-tailed, paired-samples $t$-test: $t(4999) = 262.490$, $p < 0.001$, $d = 3.713$, 95\% CI = [0.136, 0.138]; Figure S2, compare to Figure \ref{fig:fig03}A). Thus, the reduction in significant correlations following screening was unlikely to be driven solely by a reduction in statistical power.

@@

For the sake of brevity, we report here only the main pattern of findings (all results are reported in Appendix B). In the replication sample, 55 out of 393 participants (14\%) endorsed a logically invalid or improbable response on one or more of the infrequency items when completing the self-report measures. This is roughly two-thirds of the fraction of participants who were flagged for C/IE responding in the original study, suggesting that the newer quality assurance protocols used by the online platforms are at least partially effective.

In the self-report symptom measures, we replicated the finding that total scores were noticeably exaggerated in participants suspected of C/IE responding, but only for symptom measures where overall rates of symptom endorsement were the lowest (Figure S7; Table S11). Similarly, we again found that task-based screening and self-report screening measures showed low correspondence (Figure S8; Tables S12--S13); that is, excluding participants on the basis of poor behavioral performance would not have identified and removed participants who engaged in C/IE responding on self-report surveys.

Finally, when we did not apply any exclusions, we observed spurious correlations between performance on the two-step task and total scores for both symptom and personality self-report measures with a mean-shift in scores between attentive participants and participants suspected of C/IE responding (Figure S9). In contrast with our original findings, however, we found that excluding participants based on self-report \textit{or} task screening measures was sufficient to abolish these spurious correlations.

@@

To explore this possibility, we embedded attention checks into the self-report measures of two studies of patients with major depressive disorder (see Appendix C for details). Specifically, N=35 psychiatric patients (confirmed to meet criteria for a diagnosis of major depressive disorder through a structured clinical interview) across 45 unique testing sessions and N=17 healthy controls across 20 unique testing sessions, all recruited through the Rutgers-Princeton Center for Computational Cognitive Neuropsychiatry (i.e., not via online labor platforms), completed a series of self-report symptom measures, online, on their computers from the comfort of their homes. In total, 16 of 65 (24.6\%) participants failed one or more attention checks. Subdivided by group, 6 of 20 (30\%) healthy participants and 10 of 45 (22\%) psychiatric participants were flagged for C/IE responding.

Using these data, we computed pairwise Bayes factors comparing three candidate models: attention check failure rates are equal between healthy and psychiatric participants (M1); failure rates are greater in psychiatric patients (M2); and failure rates are greater in healthy participants (M3). The model assuming equal rates of failure between healthy and psychiatric participants was 2.88 times more likely than the model assuming greater rates for psychiatric patients. In turn, the model assuming lower rates of failure for psychiatric patients was 1.27 times more likely than the model assuming equal rates. Finally, the model assuming lower rates of failure for psychiatric patients was 3.65 times more likely than the model assuming higher rates for psychiatric patients. Only the final comparison exceeds the cutoff value of 3, which is conventionally treated as the minimal amount of evidence required to treat a difference in model fit as meaningful. Although the size of the sample precludes any definitive conclusion, it is noteworthy that the model least consistent with the data was the one where psychiatric patients are more likely to fail infrequency item attention checks. These data suggest, therefore, that it is unlikely that individuals with high psychiatric symptom severity were disproportionately flagged for C/IE responding in the main analyses. Accordingly, we tentatively conclude that the screening measures we are suggesting are not likely to result in overcontrol bias and false-negative correlations between tasks and symptom measures.

@@

In this study, we highlighted a particular set of circumstances, common in computational psychiatry research done on large online samples, in which spurious correlations may arise between task behavior and self-reported symptomology. When the ground-truth prevalence of a symptom is low in the general population, participants who respond carelessly on measures assessing this symptom may erroneously appear as symptomatic. Careless responding on tasks used to measure cognitive constructs can then masquerade as a correlation between individual differences in these constructs and symptom dimensions. We found repeated evidence for this pernicious pattern in two samples of participants recruited from two popular online labor platforms. False-positive rates for these spurious correlations \emph{increased} with sample size, because the correlations are due to measurement bias, not measurement noise. Importantly, we found that screening on task behavior alone was often insufficient to identify participants engaging in C/IE responding and prevent the false-positive correlations. Unfortunately, a literature review identified this type of screening as the most common practice in online computational psychiatry studies. We recommend instead to screen and exclude participants based on responding on surveys, a practice that abolished many spurious behavior-symptom correlations in our data.

@@

Viewed from the perspective of effort expenditure, our results suggest that participants appraised the cost/benefit trade-off differently for behavioral tasks and self-report surveys. Specifically, we found that only 7\% of participants in the first study were at chance-level performance in the task, compared to more than 22\% of participants who failed one or more attention-check items in the self-report surveys (a finding that qualitatively replicated in a second study involving a different task). Moreover, different measures of C/IE responding were weakly or not at all correlated between task behavior and self-report responses. This suggests the motivation for effortful responding was greater in the behavioral tasks, though precisely why is unclear. One possibility is that we gave participants a monetary incentive for attentive responding only during the tasks (a common practice, according to our literature review). A second possibility is that participants expected fewer consequences for C/IE responding during the self-report surveys, a reasonable assumption in light of how infrequently previous experiments have screened self-report data. Alternatively, participants may have found the gamified behavioral tasks more engaging or the self-report inventory more tedious. Regardless of the reason, this discrepancy reinforces our observations concerning the inadequacy of behavioral-task screening as a stand-alone method for identifying C/IE responding. Since, in general, participants may appraise costs and benefits of effortful responding differently for behavioral tasks and self-report surveys, screening for C/IE responding on one data modality may in general be unsuitable for identifying it in the other. We therefore recommend screening on each component of an experiment.
